---
title: "Assignment 3"
author: "Justin Hink"
date: "Saturday, February 14, 2015"
output: html_document
---
##Problem Set 1

The following function can be used to determine the rank of the matrix.  See comments for relevant info.

```{r}
getRank <- function(A){
  # Step 1: obrain reduced row echelon form of matrix
  
  n = nrow(A)
  m = ncol(A)
  x.pos = 1
  y.pos = 1
  
  # Tolerance for the purposes of this assignment is set here.
  tol= 1e-20 
  
  while((x.pos <= m) & (y.pos <= n))
  {
    col = A[,x.pos]
    col[1:n < y.pos] = 0
    
    # find max pivot 
    which = which.max(abs(col))
    pivot = col[which]
    
    # check for 0 pivot as defined by our tolerance
    if (abs(pivot) < tol) 
    { 
      x.pos = x.pos+1 
    } 
    else
    {
      if (which > y.pos) 
      { 
        # swap rows
        A[c(y.pos,which),]=A[c(which,y.pos),] 
      } 
        
      # pivot here
      A[y.pos,] = A[y.pos,]/pivot 
      row = A[y.pos,]
      A = A - outer(A[,x.pos],row) 
      A[y.pos,] = row
      
      y.pos = y.pos + 1
      x.pos = x.pos + 1
      
    }
  }
  
  for (i in 1:n)
    if (max(abs(A[i,1:m])) <= tol)
    {
      # move 0 rows to bottom
      A[c(i,n),] = A[c(n,i),] 
    }
  
  round(A, round(abs(log(tol,10))))
  
  # Step 2: Get non zero pivots
  
  nonZeroPivs = 0;
  for(i in 1:n)
  {
    if(A[i,i] > tol)
    {
      nonZeroPivs = nonZeroPivs + 1;
    }
  }
  
  return = nonZeroPivs
}
```

####Q1


```{r}
# Q1
A = c(1, -1, 0, 5, 2, 0, 1, 4, 3, 1, -2, -2, 4, 3, 1, -2)
dim(A) = c(4, 4)

rank = getRank(A)
rank

# check Work
library(Matrix)
checkRank = rankMatrix(A)

rank == checkRank # True
```

####Q2

For a m x n matrix where m > n, the maximum rank is n.
Generally speaking, rank(A) <= min(m,n)
Since we know that n < m, that reduces to, rank(A) <= n

The minimum rank is 1 (assuming a non zero matrix).

Rank(A) is the dimension of the image of A, i.e. the vector space consisting of Av for every vector V. 
If even one entry of A is nonzero, then there is a vector vsuch that Av is nonzero, and therefore the dimension of the image of A is not zero.

####Q3


```{r}
# Q3
A = c(1, 3, 2, 2, 6, 4, 1, 3, 2)
dim(A) = c(3, 3)

rank = getRank(A)
rank

# check work
checkRank = rankMatrix(A)

rank == checkRank # True
```
##Problem Set 2

Please see accompanying .pdf file for solution