---
title: "JHink_assign2.rmd"
author: "Justin Hink"
date: "Tuesday, February 03, 2015"
output: pdf_document
---

Problem Set 1

1) Proove A^T^A != AA^T^

Sinc in general A^T^ is not guaranteed to be equal to A, let B = A^T^

Rewrite:

BA != AB

This agrees with a fundamental property of matrices.  IE - their multiplication is not generally commutative. The proof of this theorem is widely available.

The following is a quick proof via counterexample.  We aim to falsify that A^T^A = AA^T^

```{r}
A <- matrix(seq(from=1,to=6), nrow=2, byrow=T)

```

A^T^A != AA^T^


2) A special case in which A^T^A = AA^T^ is when A is symmetrical.  In other words A^T^ = A

Problem set 2

```{r}
GetLU <- function(A){
  
  n = ncol(A)
  
  L = matrix(data=0, nrow=n, ncol=n)
  U = matrix(data=0, nrow=n, ncol=n)
  
  for(j in 1:n){
    L[j,j] = 1.0
    
    for(i in 1:j)
    {
      s1 = 0
      for(k in 1:i)
      {
        s1 = s1 + U[k,j] * L[i,k]
      }
      U[i,j] = A[i,j] - s1
    }
    
    for(i in j:n)
    {
      s2 = 0
      for(k in 1:j)
      {
        s2 = s2 +U[k,j] * L[i,k]
      }
      
      L[i,j] = (A[i,j] - s2) / U[j,j]
    }
    
    L[j,j] = 1.0
  }

  return <- list(L = L, U = U)
}



A <- c(1, 2, 1, 3, 4, 1, 5, 7, 0)
dim(A) <- c(3, 3)

result = GetLU(A)

L = result$L
U = result$U

L
U

L%*%U

```

Note that L * U is equivalent to the input matrix A

A = LU

